---
title: "DataLab R SDK"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example-usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

This document provides examples of how to use the DataLab R SDK.

```{r setup}
library(datalabSDK)
```


## Set up the client

Select the environment:

```{r}
env <- Environment$DEV
```

Authenticate and retrieve an access token:

```{r}
token <- SimplePersistedToken()$get_token(env, "temp.auth")
```

Set up the raw client:

```{r}
client <- RawClient(env)
```


## Check provenance

Provenance correlation ID before making a call: 
```{r}
client$get_provenance_session_id(token)
```

Make a call:

```{r}
search_url <- get_public_api_url(env, "search/mme")
search_results <- client$post(search_url, token, list("SearchTerm" = "sum"))
```

Provenance correlation ID after making a call:

```{r}
client$get_provenance_session_id(token)
```

## Search data

Get the correct URL for the selected environment:

```{r}
search_url <- get_public_api_url(env, "search?includeObsolete=false")
```

Define the search parameters:

```{r}
search_term <- "data"
page <- 1
max_pages <- 3
```

Perform the search:

```{r}
tryCatch(
  {
    more <- TRUE
    while(more) {
      search_body <- list(
        "SearchTerm" = search_term, 
        "Page" = page
      )
      
      search_results <- client$post(search_url, token, search_body)
      
      print(paste("First 5 results of page", page))
      
      for (result in search_results[["results"]][1:5]) {
        print(paste(result[["Name"]], result[["Description"]]))
      }
      
      page <- page + 1
      more <- search_results[["next"]] & page <= max_pages
    }
  },
  APIError = function(e) {
    stop(e)
  }
)
```

## Search models

```{r}
search_url <- get_public_api_url(env, "search/mme")

search_results <- client$post(search_url, token, list("SearchTerm" = "sum"))

print(search_results[["results"]][[1]])
```

## Upload a file

```{r}
url <- get_public_api_url(env, "")

metadata <- list(Name=jsonlite::unbox("Test file"),
                 Description=jsonlite::unbox("A test file"),
                 Tags=c("test-file"), 
                 KeyValueFields = list(),
                 AccessGroups=c("BD4BS-DEV-Public"))

data <- tempfile()
writeLines("Test file", data)

body <- list(file = httr::upload_file(data, type="text/plain"), 
             metadata = jsonlite::toJSON(metadata))

r <- client$post(url=url, token=token, body=body, encode="multipart")
```

